---
title: "Survival Analysis Report"
subtitle: "A Live Quarto Demo"
author: "Imad El Badisy"
date: today
format:
  pdf:
    toc: true
    number-sections: true
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{amsmath}
  html:
    toc: true
    theme: cosmo
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: false
  message: false
bibliography: refs.bib
---

```{r setup}
library(survival)
library(ggplot2)
library(kableExtra)
```

## Introduction

This document demonstrates how Quarto integrates **code**, **prose**,
and **mathematics** in a single reproducible file.

The dataset contains `r nrow(lung)` patients from the North Central
Cancer Treatment Group lung cancer study.
A total of `r sum(lung$status == 2, na.rm = TRUE)` deaths were
observed over the follow-up period.

We use the Cox proportional hazards model [@cox1972]:

$$
h(t \mid \mathbf{x}) = h_0(t) \exp(\mathbf{x}^\top \boldsymbol{\beta})
$$ {#eq-cox}

## Descriptive Statistics

@tbl-desc shows the baseline characteristics of the study population.

```{r}
#| label: tbl-desc
#| tbl-cap: "Baseline characteristics of lung cancer patients."

df <- lung |>
  dplyr::summarise(
    n           = dplyr::n(),
    age_mean    = round(mean(age, na.rm = TRUE), 1),
    age_sd      = round(sd(age, na.rm = TRUE), 1),
    pct_male    = round(mean(sex == 1, na.rm = TRUE) * 100, 1),
    pct_events  = round(mean(status == 2, na.rm = TRUE) * 100, 1)
  )

data.frame(
  Characteristic = c("Sample size", "Age, mean (SD)",
                     "Male sex, %", "Deaths, %"),
  Value = c(df$n,
            paste0(df$age_mean, " (", df$age_sd, ")"),
            paste0(df$pct_male, "%"),
            paste0(df$pct_events, "%"))
) |>
  kable(booktabs = TRUE, col.names = c("Characteristic", "Value")) |>
  kable_styling(latex_options = "hold_position")
```

## Kaplan-Meier Estimate

@fig-km displays the overall Kaplan-Meier survival curve.

```{r}
#| label: fig-km
#| fig-cap: "Kaplan-Meier estimate of overall survival."
#| fig-width: 6
#| fig-height: 4

km <- survfit(Surv(time, status == 2) ~ 1, data = lung)

plot(km,
     xlab = "Days", ylab = "Survival probability",
     conf.int = TRUE, col = "#0072B2", lwd = 2)
abline(h = 0.5, lty = 2, col = "grey50")
```

The median survival was **`r round(summary(km)$table["median"])`** days
(95% CI: `r round(summary(km)$table["0.95LCL"])` --
`r round(summary(km)$table["0.95UCL"])` days).

## Cox Regression

```{r}
#| label: tbl-cox
#| tbl-cap: "Cox proportional hazards model results."

fit <- coxph(Surv(time, status == 2) ~ age + sex + ph.ecog,
             data = lung)

broom::tidy(fit, exponentiate = TRUE, conf.int = TRUE) |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 3))) |>
  kable(
    booktabs = TRUE,
    col.names = c("Term", "HR", "Std. Error", "z", "p-value",
                  "95% CI Low", "95% CI High")
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

## Conclusion

Quarto renders this document to PDF and HTML from the same source.
The analysis is fully reproducible: change the data → re-render →
all numbers, figures, and tables update automatically.

## References {.unnumbered}

::: {#refs}
:::
