---
title: "Clean Quarto Report"
subtitle: "One source, HTML and PDF outputs"
author: "Imad El Badisy"
date: today
bibliography: references.bib
format:
  html:
    toc: true
    code-fold: true
    theme: cosmo
  pdf:
    toc: true
    number-sections: true
    pdf-engine: xelatex
execute:
  warning: false
  message: false
---

## Goal

This demo shows the basic Quarto workflow: write prose, run code, label outputs, cite a source, and render the same file to HTML or PDF [@quarto].

## Data

We use the built-in `mtcars` dataset so the example runs without downloads.

```{r}
#| label: tbl-summary
#| tbl-cap: "First six rows of the mtcars dataset."

knitr::kable(head(mtcars), digits = 2)
```

## Figure

Figure @fig-mpg shows the relationship between car weight and fuel economy.

```{r}
#| label: fig-mpg
#| fig-cap: "Fuel economy decreases as car weight increases."
#| echo: false

plot(mtcars$wt, mtcars$mpg,
     xlab = "Weight (1000 lbs)",
     ylab = "Miles per gallon",
     pch = 19,
     col = "#1f77b4")
abline(lm(mpg ~ wt, data = mtcars), col = "#d62728", lwd = 2)
```

## Inline Result

The dataset contains `r nrow(mtcars)` cars and `r ncol(mtcars)` variables.

## Render

```bash
quarto render report.qmd --to html
quarto render report.qmd --to pdf
```

## References

::: {#refs}
:::

