---
title: "Cross-References in Quarto"
subtitle: "Figures, Tables, Equations, and Sections"
author: "Imad El Badisy"
date: today
format:
  pdf:
    toc: true
    number-sections: true
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{amsmath}
  html:
    toc: true
    theme: flatly
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
bibliography: ../01-minimal/refs.bib
---

```{r setup}
library(survival)
library(ggplot2)
library(kableExtra)
```

## Study Design {#sec-design}

This document demonstrates Quarto's cross-referencing system.
Every figure, table, equation, and section can be referenced
by a label — and all references update automatically.

The dataset is described in @sec-data.
The statistical model is given in @eq-cox.
Results are shown in @fig-km and @tbl-cox.

## Data {#sec-data}

The `lung` dataset (NCCTG lung cancer study) contains
`r nrow(lung)` patients. See @sec-design for context.

## Statistical Model {#sec-methods}

The Cox proportional hazards model [@cox1972] is:

$$
h(t \mid \mathbf{x}) = h_0(t) \exp(\mathbf{x}^\top \boldsymbol{\beta})
$$ {#eq-cox}

The concordance index is:

$$
C = \frac
  {\sum_{i:\delta_i=1} \sum_{j:T_j > T_i}
   \mathbf{1}(\hat{r}_i > \hat{r}_j)}
  {\sum_{i:\delta_i=1} \sum_{j:T_j > T_i} 1}
$$ {#eq-cindex}

Equation @eq-cox defines the hazard; @eq-cindex defines
discrimination.

## Results

### Descriptive Statistics

@tbl-desc presents baseline characteristics.

```{r}
#| label: tbl-desc
#| tbl-cap: "Baseline characteristics of the lung cancer cohort."

data.frame(
  Variable  = c("Age, mean (SD)", "Male sex, n (%)", "Deaths, n (%)"),
  Value     = c(
    paste0(round(mean(lung$age, na.rm=TRUE), 1),
           " (", round(sd(lung$age, na.rm=TRUE), 1), ")"),
    paste0(sum(lung$sex==1, na.rm=TRUE),
           " (", round(mean(lung$sex==1, na.rm=TRUE)*100, 1), "%)"),
    paste0(sum(lung$status==2, na.rm=TRUE),
           " (", round(mean(lung$status==2, na.rm=TRUE)*100, 1), "%)")
  )
) |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = "hold_position")
```

### Survival Curves

@fig-km shows overall survival; @fig-km-sex compares by sex.

```{r}
#| label: fig-km
#| fig-cap: "Overall Kaplan-Meier survival curve."
#| fig-width: 5
#| fig-height: 3.5

km <- survfit(Surv(time, status == 2) ~ 1, data = lung)
plot(km, xlab = "Days", ylab = "Survival probability",
     col = "#0072B2", lwd = 2, conf.int = TRUE)
abline(h = 0.5, lty = 2, col = "grey50")
```

```{r}
#| label: fig-km-sex
#| fig-cap: "Kaplan-Meier curves stratified by sex."
#| fig-width: 5
#| fig-height: 3.5

km_sex <- survfit(Surv(time, status == 2) ~ sex, data = lung)
plot(km_sex,
     xlab = "Days", ylab = "Survival probability",
     col = c("#D55E00", "#0072B2"), lwd = 2,
     lty = 1:2, conf.int = FALSE)
legend("topright", legend = c("Male", "Female"),
       col = c("#D55E00", "#0072B2"), lwd = 2, lty = 1:2)
```

### Cox Regression

Results from fitting @eq-cox are in @tbl-cox.

```{r}
#| label: tbl-cox
#| tbl-cap: "Cox model: hazard ratios with 95% confidence intervals."

fit <- coxph(Surv(time, status==2) ~ age + sex + ph.ecog, data=lung)

broom::tidy(fit, exponentiate = TRUE, conf.int = TRUE) |>
  dplyr::select(term, estimate, conf.low, conf.high, p.value) |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 3))) |>
  kable(booktabs = TRUE,
        col.names = c("Variable", "HR", "95% CI Low",
                      "95% CI High", "p-value")) |>
  kable_styling(latex_options = "hold_position")
```

## Summary

Cross-references in Quarto:

| Object   | Label prefix | Reference  |
|----------|-------------|------------|
| Figure   | `fig-`      | `@fig-km`  |
| Table    | `tbl-`      | `@tbl-cox` |
| Equation | `eq-`       | `@eq-cox`  |
| Section  | `sec-`      | `@sec-methods` |

All labels update automatically when objects are added or reordered.

## References {.unnumbered}

::: {#refs}
:::
