---
title: "Survival Analysis"
subtitle: "From the Same `.qmd` Source — Now as Slides"
author: "Imad El Badisy"
date: today
format:
  beamer:
    theme: Madrid
    colortheme: seahorse
    slide-level: 2
    toc: true
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{amsmath}
        \setbeamertemplate{navigation symbols}{}
  revealjs:
    theme: moon
    slide-number: true
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
bibliography: ../01-minimal/refs.bib
---

```{r setup, include=FALSE}
library(survival)
library(ggplot2)
library(kableExtra)
```


# Background


## Motivation

- Survival data arise in clinical trials, epidemiology, economics
- Outcome: time until an event (death, relapse, failure)
- Key challenge: **censoring** — some subjects do not experience
  the event during follow-up

. . .

**Goal:** estimate the relationship between covariates and survival time.


## The Cox Model

The Cox proportional hazards model [@cox1972]:

$$
h(t \mid \mathbf{x}) = h_0(t) \exp(\mathbf{x}^\top \boldsymbol{\beta})
$$

::: {.columns}
::: {.column width="50%"}
**Advantages**

- Semi-parametric: $h_0(t)$ unspecified
- Interpretable hazard ratios
- Handles censoring naturally
:::
::: {.column width="50%"}
**Assumptions**

- Proportional hazards
- Log-linearity
- Independence
:::
:::


# Data and Methods


## Dataset

```{r}
#| echo: true

data(lung, package = "survival")
dplyr::glimpse(lung)
```


## Kaplan-Meier Curve

```{r}
#| label: fig-km
#| fig-cap: "Overall Kaplan-Meier survival estimate."
#| fig-width: 5
#| fig-height: 3.5
#| echo: false

km <- survfit(Surv(time, status == 2) ~ 1, data = lung)
plot(km, xlab = "Days", ylab = "Survival probability",
     col = "#0072B2", lwd = 2, conf.int = TRUE)
abline(h = 0.5, lty = 2, col = "grey50")
```


## Survival by Sex

```{r}
#| label: fig-km-sex
#| fig-cap: "Kaplan-Meier curves by sex."
#| fig-width: 5.5
#| fig-height: 3.5
#| echo: false

km_sex <- survfit(Surv(time, status == 2) ~ sex, data = lung)
plot(km_sex,
     xlab = "Days", ylab = "Survival probability",
     col = c("#D55E00", "#0072B2"), lwd = 2, lty = 1:2)
legend("topright", c("Male", "Female"),
       col = c("#D55E00", "#0072B2"), lwd = 2, lty = 1:2,
       bty = "n")
```


# Results


## Cox Regression Results

```{r}
#| label: tbl-cox
#| tbl-cap: "Cox model hazard ratios."
#| echo: false

fit <- coxph(Surv(time, status==2) ~ age + sex + ph.ecog, data=lung)

broom::tidy(fit, exponentiate = TRUE, conf.int = TRUE) |>
  dplyr::select(term, estimate, conf.low, conf.high, p.value) |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 3))) |>
  kable(booktabs = TRUE,
        col.names = c("Variable", "HR", "CI Low", "CI High", "p")) |>
  kable_styling(latex_options = c("hold_position", "scale_down"),
                font_size = 9)
```


## Key Findings

- Older age associated with **worse** survival (HR = 1.01 per year)
- Female sex associated with **better** survival (HR = 0.59)
- Higher ECOG score associated with **worse** survival

. . .

::: {.callout-note}
These results replicate the classic findings from the NCCTG study.
:::


# Conclusion


## Take-Home Messages

1. Quarto renders the **same `.qmd` file** to PDF, HTML, and slides
2. Code + prose stay together — no copy-paste
3. All numbers, figures, and tables update on re-render
4. Switch output format by changing one line in the YAML

. . .

```bash
quarto render slides.qmd --to beamer   # PDF slides
quarto render slides.qmd --to revealjs # HTML slides
quarto render slides.qmd --to pdf      # article PDF
```

## References {.unnumbered}

::: {#refs}
:::
